Closed-loop Neuromorphic Benchmarks

### 2015-10-14 paper/paper.tex
Updated discussion of scaling up

While the primary purpose of this paper is in describing the benchmarking
methodology, it is also worth noting that these benchmarks indicate
that the neuromorphic learning rule under investigation here is quite robost.  As
shown in Figure \ref{fig:analysis_neurons_cpu}, even just 500 neurons can consistently
adapt to control a \emph{randomly generated} 15-joint body simulation, and
deal with larger delays and noise than were seen in the example 1-joint
physical embodiment.  Since this learning system is robust across such a
wide range of conditions, and since it is efficiently implementable in a
wide variety of neuromorphic hardware, we feel it is worth further study.  This must include 
both a wider variety of minimal simulation benchmarks and also a few more 
traditional benchmarks.  These traditional
benchmarks would be particular real physical systems (specific robot arms,
for example), but testing on those would only reveal performance on those
particular arms.  As we argued here, benchmarking against a wide variety of 
randomly generated minimal simulation systems is needed to demonstrate the
space of potential situations in which neuromorphic adaptive control performs well.

@@

The approach to benchmarking presented here can be applied to a wide variety of other tasks,
by defining different minimal simulations and evaluating hardware and algorithms across a random sampling
of simulation parameters.

One possibility is to consider motor control where
the state being controlled \emph{is not the motor state itself}.  That is,
suppose we want to control the position of a hand $x$, but our output $u$
only directly controls the joints $q$ of an arm.  The position of the hand $x$
is some function of $q$, but this function may be unknown or highly complex.
Crucially, there is a learning rule that is similar to the one discussed above
that can learn this mapping (\citealt{adaptiveJacobian}), and we have had
some success in using it for particular arm control tasks (\citealt{dewolf2014}).
This would give a benchmark with a much more flexible set of capabilities;
for example a non-neural version of this learning algorithm was able to
learn to control a robot arm to follow a certain path, given only the
visual feedback of where the tip of its hand was, without knowing the
geometry of its own hand, or the position of the camera with relation to that
hand (\citealt{adaptiveJacobian}).  While it is unlikely that such a
system would work for \emph{any} possible sensorimotor combination, developing
such a benchmark could identify the space of such embodiments that a
particular algorithm running on particular neuromorphic hardware would be
good at controlling.

More generally, we believe that this approach to developing benchmarks is applicable
to entirely different kinds of tasks.  For example, much of the original work
on minimal simulation involved robot navigation and obstacle avoidance.  
\cite{Jakobi97evolutionaryrobotics} created a
minimal simulation of an environment with fixed walls, a mobile robot, and
various sensory systems.  Using such a simulation, a benchmark involving movement between locations
while avoiding collisions could be defined, and the environment could be
enhanced with moving obstacles and a richer set of sensors.  The eventual
goal might be to achieve reliable movement in an environment that includes
people.

Another example is the application of this method to classical and
operant conditioning.  Conditioning effects are extremely common in living
creatures, and are clearly evident when animals are exposed to novel environments.  As such,
it is natural to define benchmark tasks involving learning the associations between
sensory events in the environment (akin to classical conditioning) and the
associations between actions and desired sensory states (akin to operant
conditioning).  Again, the approach of minimal simulation could be used
to generate random complex sensory mappings, and various neuromorphic
learning algorithms could be evaluated in terms of their ability to
discern these patterns.  Importantly, this would also include the ability
to unlearn these patterns as the environment changes.  Interestingly, the
neuron-based classical conditioning learning rule used by \cite{verschure2003}
is quite similar to the learning rule used in the adaptive control benchmark
here.

>>>>

While the primary purpose of this paper is in describing the benchmarking
methodology, it is also worth noting that these benchmarks indicate
that the neuromorphic learning rule under investigation here is quite robust.  As
shown in Figure \ref{fig:analysis_neurons_cpu}, even just 500 neurons can consistently
adapt to control a \emph{randomly generated} 15-joint body simulation, and
deal with larger delays and noise than were seen in the example 1-joint
physical embodiment.  Since this learning system is robust across such a
wide range of conditions, and since it is efficiently implementable in a
wide variety of neuromorphic hardware, we feel it is worth further study.  This must include 
both a wider variety of minimal simulation benchmarks and also a few more 
traditional benchmarks.  These traditional
benchmarks would be particular real physical systems (specific robot arms,
for example), but testing on those would only reveal performance on those
particular arms.  As we argued here, benchmarking against a wide variety of 
randomly generated minimal simulation systems is needed to demonstrate the
space of potential situations in which neuromorphic adaptive control performs well.

@@

While the particular minimal simulation shown here suggests that
this adaptive control algorithm is worth further investigation, the overall
goal of this paper is to present the general idea of using minimal simulation
as a way to benchmark neuromorphic hardware.  That is, we believe this same
approach could be scaled up to other, more complex, closed-loop tasks.
Importantly, benchmarking these other tasks would require both the creation
of new minimal simulations \emph{and} the specification of new algorithms
suitable for performing those tasks.  These algorithms would then be
implemented with the neuromorphic hardware and connected to the minimal
simulations to construct new benchmarks.

As a first step towards scaling up, consider the more complex task of
controlling a system where the values to be controlled are not the 
joints themselves.  For example, 
suppose we want to control the position of a hand $x$, but our output $u$
only directly controls the joints $q$ of an arm.  The position of the hand $x$
is some function of $q$, but this function may be unknown or highly complex.
This is often expressed as $\dot{x}=J(q)\dot{q}$, where $J(q)$ is the
Jacobian.  In order to successfully control $x$ (the hand), the system
needs to learn the relationship $J$ that indicates how adjusting various
joints $q$ will affect the position of the hand.  Crucially, there is a
learning rule similar to the one discussed above
that can learn this mapping (\citealt{adaptiveJacobian}), and we have had
some success in using it for particular arm control tasks
(\citealt{dewolf2014}).  So far we have only tested this algorithm
in the context of one particular arm, but it was successful in learning
this relationship, and thus learning to correctly move its hand given
an unknown arm geometry.  To establish that this is a generally
useful task for neuromorphic hardware, we need to benchmark this rule against
a large family of different arms (and other systems to be controlled).  This
can be done by generating minimal simulations very similar to the
one presented here; the main difference is that there would also be a
randomly generated Jacobian function $J(q)$.  It should also be noted that
in this context, the dimensionality of $x$ and the dimensionality of $q$
are separate variables.  It may be that some algorithms work well when
$q$ is much larger than $x$, while others work best when they are similar.
Exploring this relationship is fairly straightforward with minimal simulation,
and would be an important result to know when choosing neuromorphic hardware
for a particular new situation.

Given this, we believe that the combination of minimal simulation and
neuromorphic hardware is useful for adaptive control problems in general,
whether the adaptation is in terms of an additive bias term to compensate
for external forces such as gravity (as seen in the benchmarks presented
in this paper) or if it is in terms of learning the Jacobian term relating
the controlled variables $q$ to the desired target space $x$ (as in
the adaptive Jacobian model discussed in the previous paragraph).  This
should allow systems to adapt to both unknown external forces and to unknown
bodily geometries.  However,
it is less clear whether this approach will scale to more complex robotics
tasks.

One more complex robotic task where this approach might be applicable is
navigation and obstacle avoidance.  Here, we would need both a more complex
minimal simulation for the environment, and an explicit neuromorphic
algorithm capable of performing this avoidance.  The minimal simulation itself
would need to include some sort of sensory modality (vision,
range sensing, or both), and movement in a two-dimensional environment
(probably wheeled movement, for simplicity).  To run such a simulation in
real-time, we would use many of the same tricks and simplifications used
in Jakobi's original work (\citealt{Jakobi97evolutionaryrobotics}).  These
included making separate simulations for corridors and intersections (rather
than generic simulations for any possible geometries), using noisy lookup
tables (rather than detailed physics simulations), treating collisions as
failures (rather than modelling them), and using shifting random dot patterns
for visual stimuli (rather than high-fidelity image rendering).  Given
Jakobi's success at building high-speed simulations over 20 years ago, we
believe real-time simulations of this type are feasible now.

However, having such a simulation is only half of what is required.  We would
also need a control algorithm suitable for such a situation.  This is,
itself, a topic of much research, and there is no clear best approach.
We have been exploring the use of reinforcement learning in
neural models (\citealt{Stewart2012a, Rasmussen2014b}), and note that these 
make use of the same learning algorithm as described here, with additional
neural components needed to implement action selection.  In this case, the
learning rule would adjust the system's estimate of which action is most
appropriate given the current sensory state.  We are currently investigating
this approach further.

As a more speculative possibility, we also intend to apply this approach
to tasks involving classical and operant conditioning.  Conditioning effects
are extremely common in living
creatures, and are clearly evident when animals are exposed to novel environments.  As such,
it is natural to define benchmark tasks involving learning the associations between
sensory events in the environment (akin to classical conditioning) and the
associations between actions and desired sensory states (akin to operant
conditioning).  In this case, the minimal simulations would consist of a
set of small, controlled rooms with controllable buttons and stimuli,
matching the sort of ``Skinner Box'' environments used in experimental
psychology.  The minimal simulation will also require a basic simulated
body, capable of movement, pushing buttons, and observing stimuli.  The
tasks would consist of pairing stimuli together and determining if the
learning algorithm is able to respond correctly.  For example, a model
might have a built-in response where it will salivate when presented with
food.  If the sound of a bell is paired with the presentation of food, it
should learn to salivate with presented with just the sound of a bell.
Importantly, there are extensive results showing the rate at which such
associations are learned and un-learned in various animals.  Furthermore,
we would test the ability to learn associations that are separated in time (delayed conditioning), and to recover associations that had been previously learned (spontaneous recovery).
Interestingly, there already exist
neuron-based classical conditioning learning rules that may be suitable
for such implementation, given their similarity to the learning rule used in
the adaptive control benchmark (\citealt{verschure2003}).
